Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-04-30 Documents/M1/realizableFermionicM1/sections/algebraicClosures.tex
Some adjustments to text on algebraic closures

For the two-moment model, the ME closure constructs the least biased distribution function based on the limited information available \cite{cernohorskyBludman_1994} --- namely $\cJ$ and $\vect{\cH}$.  
The ME distribution function $f_{\mbox{\tiny ME}}(\omega)$ is found by maximizing the entropy functional for particles obeying Fermi-Dirac statistics
\begin{equation}
  S[f_{\mbox{\tiny ME}}] 
  = \int_{\bbS^{2}}\big[\,(1-f_{\mbox{\tiny ME}})\log(1-f_{\mbox{\tiny ME}}) + f_{\mbox{\tiny ME}}\log f_{\mbox{\tiny ME}}\,]\,d\omega,
  \label{eq:entropyFunctional}
\end{equation} 
subject to the constraints
\begin{equation}
  \f{1}{4\pi}\int_{\bbS^{2}}f_{\mbox{\tiny ME}}(\omega)\,d\omega=\cJ
  \quad\text{and}\quad
  \f{1}{4\pi}\int_{\bbS^{2}}f_{\mbox{\tiny ME}}(\omega)\,\vect{\ell}(\omega)\,d\omega=\vect{\cH}.  
  \label{eq:closureConstraints}
\end{equation}
From the variation of the entropy functional in Eq.~\eqref{eq:entropyFunctional} with respect to $f_{\mbox{\tiny ME}}$, and introducing Lagrange multipliers to enforce the constraints in Eq.~\eqref{eq:closureConstraints}, the ME distribution takes the general form
\begin{equation}
  f_{\mbox{\tiny ME}}(\omega;a,\vect{b})=\f{1}{e^{a + \vect{b}\cdot\vect{\ell}}+1}, 
  \label{eq:fME}
\end{equation}
which can be further simplified by a change of coordinates so that $a + \vect{b}\cdot\vect{\ell}=a + b\,\mu$ ($b=|\vect{b}|$ and $\mu\in[-1,1]$).  
The Lagrange multipliers are then implicit functions of $\cJ$ and $|\vect{\cH}|$.  
The ME distribution function satisfies $0 \le f_{\mbox{\tiny ME}} \le 1$, but $a$ and $b$ are unconstrained.

Chernohorsky \& Bludman \cite{cernohorskyBludman_1994} postulate (see however \cite{lareckiBanach_2011}) that, as a function of the so-called flux saturation $h/(1-\cJ)$, the closure function $\Theta$ is independent of $\cJ$ and can be written explicitly in terms of the inverse Langevin function.  
They also provide a polynomial fit (accurate to $2\%$) given by
\begin{equation}
  \Theta_{\mbox{\tiny ME}}^{\mbox{\tiny CB}}(x)
  =x^{2}\,\big(\,3-x+3\,x^{2}\,\big).
  \label{eq:closureMECB}
\end{equation}
More recently, Larecki \& Banach \cite{lareckiBanach_2011} have shown that the explicit expression given in \cite{cernohorskyBludman_1994} is not exact, and provide another approximate expression
\begin{equation}
  \Theta_{\mbox{\tiny ME}}^{\mbox{\tiny BL}}(x)
  =\f{1}{8}\,\big(\,9\,x^{2}-5+\sqrt{33\,x^{4}-42\,x^{2}+25}\,\big),
  \label{eq:closureMEBL}
\end{equation}
which is accurate to within $0.35\%$.

>>>>

For the two-moment model, the ME closure constructs the least biased angular distribution based on the limited information available (i.e., $\cJ$ and $\vect{\cH}$) \cite{cernohorskyBludman_1994}.  
The ME distribution $f_{\mbox{\tiny ME}}(\omega)$ is found by maximizing the entropy functional, which for particles obeying Fermi-Dirac statistics is given by
\begin{equation}
  S[f_{\mbox{\tiny ME}}] 
  = \int_{\bbS^{2}}\big[\,(1-f_{\mbox{\tiny ME}})\log(1-f_{\mbox{\tiny ME}}) + f_{\mbox{\tiny ME}}\log f_{\mbox{\tiny ME}}\,]\,d\omega,
  \label{eq:entropyFunctional}
\end{equation} 
subject to the constraints
\begin{equation}
  \f{1}{4\pi}\int_{\bbS^{2}}f_{\mbox{\tiny ME}}(\omega)\,d\omega=\cJ
  \quad\text{and}\quad
  \f{1}{4\pi}\int_{\bbS^{2}}f_{\mbox{\tiny ME}}(\omega)\,\vect{\ell}(\omega)\,d\omega=\vect{\cH}.  
  \label{eq:closureConstraints}
\end{equation}
From the variation of the entropy functional in Eq.~\eqref{eq:entropyFunctional} with respect to $f_{\mbox{\tiny ME}}$, and introducing Lagrange multipliers to enforce the constraints in Eq.~\eqref{eq:closureConstraints}, the ME distribution takes the general form
\begin{equation}
  f_{\mbox{\tiny ME}}(\omega;a,\vect{b})=\f{1}{e^{a + \vect{b}\cdot\vect{\ell}}+1}, 
  \label{eq:fME}
\end{equation}
which can be further simplified by a change of coordinates $a + \vect{b}\cdot\vect{\ell}=a + b\,\mu$, where $b=|\vect{b}|$, and $\mu\in[-1,1]$ is the cosine of the angle between $\vect{b}$ and $\vect{\ell}$.  
The Lagrange multipliers are implicit functions of $\cJ$ and $|\vect{\cH}|$.  
The ME distribution function satisfies $0 \le f_{\mbox{\tiny ME}} \le 1$, but $a$ and $b$ are unconstrained.

Chernohorsky \& Bludman \cite{cernohorskyBludman_1994} postulate (but see \cite{lareckiBanach_2011}) that, as a function of the so-called flux saturation $x=h/(1-\cJ)$, the closure function $\Theta(x)$ is independent of $\cJ$ and can be written explicitly in terms of the inverse Langevin function.  
They also provide a polynomial fit (accurate to $2\%$) given by
\begin{equation}
  \Theta_{\mbox{\tiny ME}}^{\mbox{\tiny CB}}(x)
  =x^{2}\,\big(\,3-x+3\,x^{2}\,\big).
  \label{eq:closureMECB}
\end{equation}
More recently, Larecki \& Banach \cite{lareckiBanach_2011} have shown that the explicit expression given in \cite{cernohorskyBludman_1994} is not exact, and provide another approximate expression
\begin{equation}
  \Theta_{\mbox{\tiny ME}}^{\mbox{\tiny BL}}(x)
  =\f{1}{8}\,\big(\,9\,x^{2}-5+\sqrt{33\,x^{4}-42\,x^{2}+25}\,\big),
  \label{eq:closureMEBL}
\end{equation}
which is accurate to within $0.35\%$.  
On the interval $x\in[0,1]$, the curves given by Eqs.~\eqref{eq:closureMECB} and \eqref{eq:closureMEBL} lie practically on top of each other.
